---
title: "Hands-on_Exercise01"
author: "guyue"
---

1.0 **Installing and loading the required libraries**

```{r}
pacman::p_load(tidyverse)
```

1.1 **Importing data (csv)**

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

1.2 Check the missing data

```{r}
exam_data %>%
  summarise(
    across(everything(), ~ sum(is.na(.)))
  )
```

2.0 ggplot2-plot (geom)

2.1 pointplot

2.1.1 aes(color , size , shape, alpha )

```{r}
ggplot(exam_data,
       aes(MATHS,
           ENGLISH))+
  geom_point(aes(
    color=CLASS))
```

2.1.2 geom_smooth (fit line)

```{r}
ggplot(exam_data,
       aes(MATHS,
           ENGLISH))+
  geom_point(aes(color=CLASS))+
  geom_smooth(method=lm, 
              size=0.5)
```

2.2 bar-plot

for discrete data

```{r}
ggplot(exam_data,
       aes(CLASS))+
  geom_bar()
         
```

flip

```{r}
ggplot(exam_data,
       aes(CLASS))+
  geom_bar()+
  coord_flip()
```

2.3 histogram

for continue data

```{r}
ggplot(exam_data,
       aes(MATHS))+
  geom_histogram()
```

geom(bin width)

```{r}
ggplot(exam_data,
       aes(MATHS))+
  geom_histogram()
geom_histogram(binwidth = 5)
```

ggplot (fill)

```{r}
ggplot(exam_data,
       aes(MATHS,
           fill=CLASS))+
  geom_histogram()
geom_histogram(binwidth = 1)
```

geom (position)

```{r}
ggplot(exam_data,
       aes(MATHS,
           fill=CLASS))+
  geom_histogram()+
  geom_histogram(binwidth = 5,
               position = "identity")
```

2.4 density-plot

```{r}
ggplot(exam_data,
       aes(MATHS,
           color=RACE))+
  geom_density()
```

geom (alpha) ??????

```{r}
ggplot(exam_data,
       aes(MATHS,
           color=RACE))+
  geom_density(alpha=0.2)
```

2.5 box-plot

```{r}
ggplot(exam_data,
       aes(MATHS))+
  geom_boxplot()
```

more variables

```{r}
ggplot(exam_data,
       aes(MATHS,
           CLASS))+
  geom_boxplot()
```

2.6 point-plot

```{r}
ggplot(exam_data,
       aes(MATHS,
           ENGLISH))+
  geom_point()
```

2.7 **山峦图**

ggridges package install

```{r}
pacman::p_load("ggridges")
```

mountain plot

```{r}
ggplot(exam_data,
       aes(
    x = MATHS,
    y = RACE,
    fill = CLASS
  )) +
  ggridges::geom_density_ridges()
```

3.0 ggplot2-plot (coordinate)

x,y label

```{r}
ggplot(exam_data,
       aes(MATHS,
           ENGLISH))+
  geom_point(aes(
    color=CLASS))+
  scale_x_continuous("maths-scores")+
  scale_y_continuous("english-scores")

```

坐标轴限制: coord_cartesian(xlim=c(0,100), ylim=c(0,100))

```{r}
ggplot(exam_data,
       aes(MATHS,
           ENGLISH))+
  geom_point(aes(
    color=CLASS))+
  scale_x_continuous("maths-scores")+
  scale_y_continuous("english-scores")+
  coord_cartesian(xlim=c(0,100), ylim=c(0,100))
```
