---
title: "Hands-on_Exercise09.html"
---

# Hands-on Exercise 10

Information Dashboard Design: R methods

PUBLISHED

March 21, 2024

MODIFIED

March 21, 2024

## 1 **Overview**

In this exercise I will:

-   create bullet chart by using ggplot2,

-   create sparklines by using ggplot2 ,

-   build industry standard dashboard by using R Shiny.

## 2 **Loading libraries**

```{r}
pacman::p_load(lubridate, ggthemes, reactable,
reactablefmtr, gt, gtExtras, tidyverse)
```

## 3 The data set

### 3.1 **Importing data**

```{r}
coffeechain <- read_rds("data/rds/CoffeeChain.rds")
```

### 3.2 Data preparation

Aggregate Sales and Budgeted Sales at the Product level.

```{r}
product <- coffeechain %>%
  group_by(`Product`) %>%
  summarise(`target` = sum(`Budget Sales`),
            `current` = sum(`Sales`)) %>%
  ungroup()
```

### 3.3 **Bullet chart in ggplot2**

```{r}
ggplot(product, aes(Product, current)) + 
  geom_col(aes(Product, max(target) * 1.01),
           fill="grey85", width=0.85) +
  geom_col(aes(Product, target * 0.75),
           fill="grey60", width=0.85) +
  geom_col(aes(Product, target * 0.5),
           fill="grey50", width=0.85) +
  geom_col(aes(Product, current), 
           width=0.35,
           fill = "black") + 
  geom_errorbar(aes(y = target,
                    x = Product, 
                    ymin = target,
                    ymax= target), 
                width = .4,
                colour = "red",
                size = 1) +
  coord_flip()
```

## 
